A Note on Convergence Diagnostics for Multiple Imputation using Chained Equations (MICE)

### 2019-11-12 Simulation/Writeup/preprint.tex
Decided to split the Research Report and the final article (Thesis). RR is in Writeup folder of Simulation now.

Moreover, $\widehat{R}$ could  theoretically not be smaller than one, yet it happened several times in this study (see online appendix XYZ). How could $\widehat{R}$ smaller than 1 occur? The number of simulations is smaller than in 'regular' MCMC processes. Therefore, the '$(n-1/n)$' correction factor can influence the estimated potential scale reduction factor. This downwards bias is in the opposite direction than expected: "The mixture-of-sequences variance, V, should stabilize as a function of n. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)" \cite[p~438]{broo98}.

@@

This note illustrates that conventional convergence diagnostics behave differently on MI data as compared to the output of 'regular' MCMC algorithms. The recommended threshold for $\widehat{R}$ is too stringent for MI data.

@@

The results in this paper were obtained using \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the \pkg{mice}~\Sexpr{paste(packageVersion("mice"), collapse = ".")} package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

>>>>

Moreover, $\widehat{R}$ could  theoretically not be smaller than one, yet it happened several times in this study (see online appendix XYZ). How could $\widehat{R}$ smaller than 1 occur? The number of simulations is smaller than in 'regular' MCMC processes. Therefore, the '$(n-1/n)$' correction factor can influence the estimated potential scale reduction factor. This downwards bias is in the opposite direction than expected.

@@

"The mixture-of-sequences variance, V, should stabilize as a function of n. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)" \cite[p~438]{broo98}.

@@

The results in this paper were obtained using \proglang{R}~3.6.1 with the \pkg{mice}~3.6.0.9000 package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.
